---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Validity

<!-- badges: start -->
<!-- badges: end -->

The goal of the "Validity" package is to implement the method of the paper "A Test for the Validity of Regression Models". 

Citation: 
*Frahm, Gabriel, A Test for the Validity of Regression Models (October 23, 2023). Available on SSRN:*
<https://ssrn.com/abstract=4610329> or <http://dx.doi.org/10.2139/ssrn.4610329>

## Disclaimer

This package is a pre-alpha version. The execution time of the “validity” function in particular is still in need of improvement, especially if the aim is to run a comprehensive simulation. At present, it takes about 2.5 seconds to perform the validity test with n=100. With n>=1000 or more than one regressor you also need a lot of RAM. In these cases, it is recommended to adjust the default bootstrap of N=1000 in the "validity" function (e.g. to N=100). Especially since R uses only one CPU core in standard mode. 

Comments, help, and suggestions for improvement are greatly appreciated at [schuetze@hsu-hh.de](schuetze@hsu-hh.de).

## Installation

You can install the development version of Validity from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("floschuetze/Validity")
```

## Primary Function

The "validity" function represents the fundamental component of the package. It performs a statistical test with the objective of determining the validity of a regression model that requires a dependent variable, which is represented as a vector of length "n". Additionally, it requires regressors, which could be a vector or a matrix of length "m*n". By default, an OLS regression is run using a constant and all regressors to explain y. 

For additional input options, please refer to the help file for the validity function. This is particularly the case when a regression model other than ordinary least squares (OLS) is to be employed.

The following example illustrates the application of the validity function in the context of a standard OLS regression, whereby the null hypothesis (H0) is rejected.

```{r example2}
library(Validity)
#It is evident that the model "y=constant+beta*x" is not valid.
x <- rnorm(100, mean=0, sd=1)
y <- x^3-x^2 # true model
validity(y,x) 
```


## Additional optional functions within the package

### Default Regression function
  
It is possible to specify your own regression function and pass it to the "validity" function. It is important that your custom function has as output the estimated values based on the custom regression for y as a vector. 

The following regression function is the default simple OLS regression used in the function "validity".

```{r reg, eval=FALSE}
reg<-function(y, X) {
  d<-data.frame(X,y1=y)
  model <- lm(y1 ~ ., data = d)
  return(as.vector(model$fitted.values))
}
```


### Example 1

The following function can be used to replicate the first example from the "A Test for the Validity of Regression Models" paper. 

```{r function1, eval=FALSE}
#c and n must be specified; the function returns the p-value; tau is optional
Example1 <- function(c,n,tau) {
  if (missing(tau)) tau <- 1
  x <- rnorm(n, mean=0, sd=1)
  y <- rnorm(n, mean=0, sd=tau)
  y[which(x>0)]<- y[which(x>0)]+c
  y[which(x<=0)]<- y[which(x<=0)]-c
  p<-validity(y,x,text=0)$p_value
  return(p)
}
```

### Example 2

The following function can be used to replicate the second example from the "A Test for the Validity of Regression Models" paper. 

```{r function2, eval=FALSE}
#c and n must be specified; the function returns the p-value; tau is optional
Example2 <- function(c,n,tau) {
  if (missing(tau)) tau <- 1
  x <- rnorm(n, mean=0, sd=1)
  y<-(-1)+x+c*((x^2)-1)+rnorm(n, mean=0, sd=tau)
  p<-validity(y,x,text=0)$p_value
  return(p)
}
```

### Example 3

The following function can be used to replicate the third example from the "A Test for the Validity of Regression Models" paper. 

```{r function3, eval=FALSE}
#c and n must be specified; the function returns the p-value; tau is optional
Example3 <- function(c,n,tau) {
  if (missing(tau)) tau <- 1
  require(MASS)
  sigma<-rbind(c(1,0.5), c(0.5,1))
  mu<-c(0, 0) 
  LK<-as.matrix(mvrnorm(n=n, mu=mu, Sigma=sigma))
  y<-0.25*LK[,1]+0.75*LK[,2]+c*LK[,1]*LK[,2]+rnorm(n, mean=0, sd=tau)
  p<-validity(y,LK,text=0)$p_value
  return(p)
}
```
